---
title: "Getting started with heatindexbr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with heatindexbr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

`heatindexbr` provides spatially interpolated Heat Index (Rothfusz 1990)
rasters at ~1 km resolution for the 1,477 municipalities of Brazil's official
Semiarid region (CONDEL/SUDENE Resolution 176/2024). The main product is a
climatology for 2000-2025 covering all 12 months and 24 local hours. This
vignette walks through the core workflow: search, download, extract, and plot.

## Installation

```{r install, eval=FALSE}
remotes::install_github("antcneto/heatindexbr")
```

```{r setup}
library(heatindexbr)
```

## What data is available?

Use `hi_available()` to see the products, temporal coverage, and available
months and hours:

```{r available}
hi_available()
```

## Searching for municipalities

Municipality names follow the IBGE convention: **uppercase without accents**.
Use `hi_search()` to find the correct name before calling other functions:

```{r search}
hi_search("Mossoro", state = "RN")
hi_search("Campina Grande", state = "PB")
```

> **Note:** always use the `name_muni` value returned by `hi_search()` in
> subsequent calls. For example, use `"MOSSORO"`, not `"Mossoró"`. This
> limitation follows the IBGE naming standard.

## Extracting Heat Index for a municipality

`hi_municipality()` downloads the municipal table on first use (~29 MB,
cached locally) and returns summary statistics with NOAA classification.
No raster download is required for this operation.

```{r municipality}
result <- hi_municipality("MOSSORO", state = "RN",
                          month = 1, hour_local = 15)
result
```

To retrieve all months and hours for a municipality:

```{r municipality-all, eval=FALSE}
hi_municipality("MOSSORO", state = "RN")
```

The `noaa_class` column uses the standard NOAA Heat Index thresholds:

| Class | IC range | Risk |
|---|---|---|
| No caution | < 27°C | Low |
| Caution | 27-32°C | Fatigue with prolonged exposure |
| Extreme caution | 32-41°C | Heat cramps or exhaustion possible |
| Danger | 41-54°C | Heat exhaustion likely |
| Extreme danger | > 54°C | Heat stroke imminent |

## Downloading and plotting the raster

`hi_download()` downloads the full 288-band climatology stack (~1 GB) on
first use and caches it locally. Subsequent calls read from the cache and
extract only the requested bands in memory.

```{r download, eval=FALSE}
# Single band: January at 15h local time
r <- hi_download(month = 1, hour_local = 15)

# All hours for July (24-band raster)
r_july <- hi_download(month = 7)

# Full 288-band stack
r_stack <- hi_download()
```

```{r plot, eval=FALSE}
hi_plot(r, title = "Heat Index — January, 15h local")
```

> **Note on hours:** all hours are specified in local time (UTC-3). The
> conversion to UTC is handled internally. For example, `hour_local = 15`
> retrieves the band stored at 18h UTC.

## Extracting for a custom polygon

Use `hi_shape()` with any `sf` polygon object:

```{r shape, eval=FALSE}
library(sf)
my_area <- st_read("my_area.gpkg")
result  <- hi_shape(my_area, month = 1, hour_local = 15)
result$stats
result$raster
```

## Working with the 2025 synoptic product

The original 2025 product (four synoptic hours, annual means) remains
available via the `product` argument:

```{r synoptic, eval=FALSE}
r_2025 <- hi_download(product = "synoptic_2025", hour_local = 15)
hi_municipality("MOSSORO", state = "RN",
                product = "synoptic_2025", hour_local = 15)
```

## Citation

```{r citation, eval=FALSE}
citation("heatindexbr")
```

Climatology dataset (2000-2025):

> Campos Neto, A. (2026). *Heat Index climatology for the Brazilian Semiarid
> (2000-2025): 1 km rasters, municipal table, and population exposure data*.
> Zenodo. https://doi.org/10.5281/zenodo.21049066

Synoptic 2025 dataset:

> Campos Neto, A. (2025). *heatindexbr: Heat Index rasters for Brazil's
> Semiarid region* (v0.1.0). Zenodo.
> https://doi.org/10.5281/zenodo.20942619
