---
title: "Getting Started with heatindexbr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with heatindexbr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

## Introduction

`heatindexbr` provides pre-computed Heat Index (HI) rasters for the
**Brazilian Semiarid region** (1,477 municipalities, Res. CONDEL/SUDENE
176/2024) at four synoptic hours and tools for spatial extraction.

The Heat Index follows the Rothfusz (1990) formulation adopted by the
US National Weather Service and combines air temperature and relative
humidity into a single thermal stress metric.

## Installation

```{r install}
# From GitHub (development version)
# install.packages("remotes")
remotes::install_github("antcamposneto/heatindexbr")
```

## Basic usage

### Search for a municipality

```{r search}
library(heatindexbr)

# Find municipalities named "Mossoró"
hi_search("Mossoro")
#>   code_muni   name_muni abbrev_state
#> 1   2408102     MOSSORO           RN
```

### Extract Heat Index for a municipality

```{r extract}
# Heat Index at 15h local time (peak heat stress)
result <- hi_municipality("Mossoro", state = "RN", hour = "15h")
print(result)
#>   code_muni name_muni abbrev_state     mean    noaa_class hour year
#> 1   2408102   MOSSORO           RN 33.47814 Extreme caution  15h 2025
```

### Extract for multiple municipalities

```{r multiple}
nordeste_capitals <- c(2408102, 2304400, 2611606, 2507507)
hi_municipality(nordeste_capitals, hour = "15h",
                stat = c("mean","min","max"))
```

### Extract for a custom shapefile

```{r shape}
library(sf)
my_area <- st_read("minha_area_de_estudo.shp")
result  <- hi_shape(my_area, hour = "15h", return_raster = TRUE)

# Plot the clipped raster
hi_plot(result$raster, title = "Heat Index — Study Area")
```

### All four synoptic hours

```{r cycle}
# Download all hours
paths <- hi_download(year = 2025, hours = c("00h","09h","15h","21h"))

# Extract for one municipality across all hours
hours <- c("00h","09h","15h","21h")
ciclo <- lapply(hours, function(h) {
  hi_municipality("Cabrobo", state="PE", hour=h)
}) |> do.call(rbind, args=_)

print(ciclo[, c("name_muni","hour","mean","noaa_class")])
```

## NOAA Classification

| Class | IC range |
|-------|----------|
| No caution | < 27°C |
| Caution | 27–32°C |
| Extreme caution | 32–41°C |
| Danger | 41–54°C |
| Extreme danger | > 54°C |

## Citation

```{r citation}
citation("heatindexbr")
```

Campos Neto, A. et al. (2026). Heat Index mapping for the Brazilian Semiarid:
spatial interpolation under sparse monitoring networks.
*International Journal of Climatology*. [In review]
