---
title: "Getting started with heatindexbr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with heatindexbr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

`heatindexbr` provides spatially interpolated Heat Index (Rothfusz 1990) rasters
at ~1 km resolution for the 1,477 municipalities of Brazil's official Semiarid
region (CONDEL/SUDENE Resolution 176/2024). This vignette walks through the
core workflow: search → download → extract → plot.

## Installation

```{r install, eval=FALSE}
# Install from GitHub
remotes::install_github("antcneto/heatindexbr")
```

```{r setup}
library(heatindexbr)
```

## What data is available?

Use `hi_available()` to see the years, resolutions, and synoptic hours
currently distributed:

```{r available}
hi_available()
```

## Searching for municipalities

Municipality names in `heatindexbr` follow the IBGE convention: **uppercase
without accents**. Use `hi_search()` to find the correct name:

```{r search}
# Case-insensitive partial match
hi_search("Mossoro", state = "RN")
hi_search("Campina Grande", state = "PB")
```

> **Important:** Always use the `name_muni` value returned by `hi_search()`
> when calling `hi_municipality()`. For example, `"MOSSORO"`, not `"Mossoró"`.

## Extracting Heat Index for a municipality

`hi_municipality()` downloads the raster (cached on first use), crops it to
the municipality boundary, and returns summary statistics with NOAA
classification:

```{r municipality}
result <- hi_municipality("MOSSORO", state = "RN", hour = "15h")
result
```

The `noaa_class` column uses the standard NOAA Heat Index thresholds:

| Class | IC range | Risk |
|---|---|---|
| No caution | < 27°C | Low |
| Caution | 27–32°C | Fatigue with prolonged exposure |
| Extreme caution | 32–41°C | Heat cramps or exhaustion possible |
| Danger | 41–54°C | Heat exhaustion likely |
| Extreme danger | > 54°C | Heat stroke imminent |

You can retrieve all four synoptic hours at once by leaving `hour = NULL`:

```{r all-hours, eval=FALSE}
hi_municipality("MOSSORO", state = "RN", hour = NULL)
```

## Downloading and plotting the full raster

`hi_download()` returns a `SpatRaster` (terra package). The raster is
downloaded once and cached in your user data directory:

```{r download}
r <- hi_download(year = 2025, hour = "15h")
r
```

Pass the raster directly to `hi_plot()` for a ready-to-use map:

```{r plot}
hi_plot(r,
        title = "Annual mean Heat Index — 15h (2025)",
        hour  = "15h",
        year  = 2025)
```

## Extracting for a custom polygon

If you have your own `sf` object (e.g. a watershed, a state boundary, or a
study area), use `hi_shape()`:

```{r shape, eval=FALSE}
library(sf)

# Example: using your own polygon
my_area <- st_read("my_area.gpkg")

result <- hi_shape(my_area, hour = "15h")
result$stats   # data.frame with mean, min, max, noaa_class
result$raster  # clipped SpatRaster
```

## Technical note on the raster

The underlying raster (`IC_2025_annual_all.tif`) is stored in **EPSG:5880**
(SIRGAS 2000 / Brazil Polyconic) — the appropriate projected CRS for
Brazil. Do not use the WGS84 version, which has a clipping artifact on the
eastern boundary. See `vignette("methodology")` for full details on the
interpolation method.

## Citation

If you use `heatindexbr` in your research, please cite:

```{r citation, eval=FALSE}
citation("heatindexbr")
```

Or cite the Zenodo dataset directly:

> Campos Neto, A. (2025). *heatindexbr: Heat Index rasters for Brazil's
> Semiarid region* (v0.1.0). Zenodo.
> https://doi.org/10.5281/zenodo.20942619
